---
title: "Quiz 04"
author: ""
date: "Nov 09, 2021"
output: 
  html_notebook:
    theme: spacelab
    toc: yes
    toc_float: yes
---

<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r echo = FALSE}
######################################################################
##################### FOLLOW THESE DIRECTIONS ########################
### 1. Don't delete this code chunk, it won't appear in the output ###
### 2. Put your name in the author field above                     ###
### 3. Write the code for all questions in this R Notebook         ###
### 4. ALWAYS comment your code.                                   ###
### 5. Press the knit button when you're done, your code must      ###
###    run with no errors in order for this to work. If your       ###
###    code has errors, comment them out                           ###
######################################################################
######################################################################
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", tidy = FALSE, max.print = 20)
```

- Read in the saccharomyces_cerevisiae.fasta and saccharomyces_cerevisiae.gff to answer the questions 
```{r}
.libPaths('/projects/community/py-image/2019.10A/bd387/anaconda3-intel17/envs/shahclass/lib/R/library')
library(rtracklayer)
library(Biostrings)
library(tidyverse)

yeast_gff <- readGFFAsGRanges("saccharomyces_cerevisiae.gff")
yeast_fasta <- readDNAStringSet("saccharomyces_cerevisiae.fasta")
```

1. Using the genome annotation file, create a density plot of gene lengths of all `tRNA_gene`. (2 pts)
2. Create a DNAStringSet of all `tRNA_gene` sequences of length 72 bp. Be aware of strand orientation. (3 pts)

Use the code from Lab-16 to create a DNAStringSet of CDS of all intron-less genes in yeast.

3. How many genes have stretches of 4 guanines (GGGG), and which gene has the most number of these stretches? Count overlapping stretches as well. For instance GGGGG = 2 GGGG stretches (3 pts)
4. Starting with CDS of all intron-less genes in yeast, make a scatterplot of the ratio of Lys/Arg amino acids on the y-axis and CDS length of genes on the x-axis. Appropriately transform the axes. (2 pts)

