---
title: "Lab 07 - homework solutions"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 4
    toc_float: TRUE
---

### Packages
```{r}
library(rtracklayer)
library(GenomicRanges)
```

#### 1. How many protein-coding genes are on the mitochondria?
```{r}
yeast.gr <- readGFFAsGRanges("~/Google Drive/rutgers/2018_S/lab07/saccharomyces_cerevisiae.gff")
yeast.gr

# Method 1
mt.gr <- yeast.gr[seqnames(yeast.gr)=="chrmt"]
genes.mt.gr <- mt.gr[mt.gr$type=="gene"]
genes.mt.gr

# Method 2
genes.gr <- yeast.gr[yeast.gr$type=="gene"]
genes.mt.gr2 <- genes.gr[seqnames(genes.gr)=="chrmt"]
genes.mt.gr2

# Method 3
genes.mt.gr3 <- yeast.gr[seqnames(yeast.gr)=="chrmt" & yeast.gr$type=="gene"]
genes.mt.gr3
```

#### 2. Which chromosome has the most number of protein-coding genes beginning with “YA”? (+1 Bonus)
```{r}
genes.gr <- yeast.gr[yeast.gr$type=="gene"]
list.of.first2char <- substr(genes.gr$Name, start = 1, stop = 2)

genes.YA.gr <- genes.gr[list.of.first2char=="YA"]
table(seqnames(genes.YA.gr))
```

#### 3. Which chromosome has the most number of tRNA genes?
```{r}
tRNA.gr <- yeast.gr[yeast.gr$type == "tRNA_gene"]
table(seqnames(tRNA.gr))
```

#### 4. Which chromosome has the longest protein-coding gene?
```{r}
gene.lengths <- width(genes.gr)
max(gene.lengths)

which(gene.lengths == max(gene.lengths))
which.max(gene.lengths)
genes.gr[which.max(gene.lengths)]

as.character(seqnames(genes.gr[which.max(gene.lengths)]))
```

#### 5. Find the longest gene on the smallest nuclear chromosome.
```{r}
chromosome.gr <- yeast.gr[yeast.gr$type=="chromosome" & seqnames(yeast.gr)!="chrmt"]

# Alternate way to get nuclear chromosomes
nuclear.gr <- dropSeqlevels(yeast.gr, value = "chrmt", pruning.mode = "coarse")
nuclear.chromosomes.gr <- nuclear.gr[nuclear.gr$type=="chromosome"]

chrom.length <- width(nuclear.chromosomes.gr)
nuclear.chromosomes.gr[which(chrom.length == min(chrom.length))]

# Chr I is the smallest chromosome

chrI.genes.gr <- genes.gr[seqnames(genes.gr)=="chrI"]
chrI.genes.gr2 <- keepSeqlevels(genes.gr, "chrI", pruning.mode = "coarse")

gene.lengths.chrI <- width(chrI.genes.gr)
which.max(gene.lengths.chrI)

chrI.genes.gr[which.max(gene.lengths.chrI)]
```

#### 6. How many protein-coding genes have a hyphen in their name? Use Name column.
```{r}
# Method 1
genenames <- genes.gr$Name
split.genenames <- strsplit(genenames, "-")

split.lengths <- c()

for(counter in 1:length(split.genenames)){
  split.lengths[counter] <- length(split.genenames[[counter]])
}

table(split.lengths)
sum(split.lengths>1)

# Method 2
genenames <- genes.gr$Name
grep.output <- grep("-", genenames)
length(grep.output)
```

#### 7. Find the second longest gene on the negative strand. (+1 Bonus)
```{r}
negative.strand.genes.gr <- genes.gr[strand(genes.gr)=="-"]
gene.lengths.neg <- width(negative.strand.genes.gr)

which.max(gene.lengths.neg)

negative.strand.genes.without.max.gr <- negative.strand.genes.gr[-which.max(gene.lengths.neg)]

which.max(width(negative.strand.genes.without.max.gr))

negative.strand.genes.without.max.gr[1958]
```






