---
title: "Quiz 05 solutions"
output: html_notebook
---

```{r}
library(tidyverse)
library(MASS)
library(broom)
```


1. Find the correlation between x and y in data1.csv. Which correlation coefficient is most appropriate here and why? (2 pts)
```{r}
data1 <- read_csv("data1.csv")
data1
```
```{r}
ggplot(data1, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm")
```
```{r}
cor(data1$x, data1$y, method = "spearman")
cor(data1$x, data1$y, method = "pearson")

```

```{r}
ggplot(data1, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_y_continuous(trans = "log10")
```


2. What test would be appropriate to compare x and y in data2.csv and why? Do you find similar results when you use other statistical tests? (2 pts)
```{r}
data2 <- read_csv("data2.csv")
data2 %>%
  summary()
```
```{r}
ggplot(data2, aes(x = x)) +
  geom_density()

ggplot(data2, aes(x = y)) +
  geom_density()

ggplot(data2, aes(x = x)) +
  geom_density(col = "red") +
  geom_density(aes(y), col = "black")
```

```{r}
t.test(data2$x, data2$y)
wilcox.test(data2$x, data2$y)
ks.test(data2$x, data2$y)
```


3. You’ve performed differential gene expression analyses on a sample. The data is in data3.csv. How many genes can be classified as significantly different at a false discovery rate of 0.05? (2 pts)
```{r}
data3 <- read_csv("data3.csv")
data3
```

```{r}
data3 <- data3 %>%
  mutate(padjust = p.adjust(pvalue, method = "fdr"))
```

```{r}
data3 %>%
  filter(padjust < 0.05) %>%
  nrow()

data3 %>%
  mutate(signif = ifelse(padjust<0.05, TRUE, FALSE)) %>%
  pull(signif) %>%
  sum()
```


4. Is there a difference in the log2FoldChange.rna between genes that are significantly different and the ones that are not? Perform appropriate statistical test. (1 pt)
```{r}
ggplot(data3, aes(x = log2FoldChange.rna, color = padjust < 0.05)) +
  geom_density()
```

```{r}
data3.sig <- data3 %>%
  filter(padjust < 0.05)
data3.nonsig <- data3 %>%
  filter(padjust >= 0.05)

t.test(data3.sig$log2FoldChange.rna, data3.nonsig$log2FoldChange.rna)
wilcox.test(data3.sig$log2FoldChange.rna, data3.nonsig$log2FoldChange.rna)
```


5. Using human_features.csv, identify features that best predict the CAI values of genes. Which feature has the biggest effect on predicting CAI values? Perform appropriate transformations of the variables. (3 pt)
```{r}
human <- read_csv("human_features.csv")
human

summary(human)
```

```{r}
human[1, ]
```

```{r}
human %>%
  gather(Variables, Value, 3:11) %>%
  ggplot(., aes(x = Value)) +
    geom_density() +
    facet_wrap(~Variables, scales = "free")
```

```{r}
cai.lm <- lm(CAI ~ CDS_GC_Content + Transcript_FE + Translate_FE + UTR3_GC_Content +
               UTR5_GC_Content + log10(CDS_Length) + log10(UTR3_Length + 1) + log10(UTR5_Length + 1) + log10(uATG_count + 0.1), data = human)

cai.lm %>% tidy()
cai.lm %>% glance()
```


```{r}
stepAIC(cai.lm)
```















