---
title: "Lab 08 - Loops, functions, and lists"
author: "Premal Shah"
date: "Oct 05, 2021"
output: 
  html_notebook:
    theme: spacelab
    toc: yes
    toc_float: yes
---

## Homework from Lab-07

1. Write a `for` loop to calculate the length of a vector (you cannot use the function `length` anywhere). (2 pts)
2. Write a `for` loop to reverse the order in a vector. For example, if you start with (1,2,3,4,5) it should return (5,4,3,2,1). (3 pts)
3. Write a nested `for` loop to calculate sums of rows of a matrix. This is equivalent to `rowsums`. You cannot use `sum` function. (5 pts)

Bonus pts
1. Write a nested `for` loop to return the first 10 numbers of the Fibonacci series. (3 pts)

## Indefinite loops
- `while` loop
```{r eval = FALSE}
while(condition is TRUE){
  do something 1
  do something 2
  ...
}
```

## Factorial function
```{r}
n <- 13

fact_var <- 1
while(n >= 1){ # While n is greater than 1 keep going through the loop
  fact_var <- fact_var*n # Multiply n to the temporary factorial variable
  n <- n - 1 # Reduce the value of n by 1
}
fact_var
```


## Exercise 8.1
1. Keep rolling an unbiased dice until you see a 1. How many dice rolls did it take?
2. You have a biased dice where even numbers are 3 times as likely as odd numbers. Keep rolling it until you see a 5. 
3. Stop rolling the above dice till you see a 1 or once you have rolled the dice 10 times, whichever comes first.

## Custom functions

## Lists
- Comparisons with data frames

### Accessing and subsetting Lists
- Like a data frame
- Using indices

## Apply functions
 - `apply`
 - `lapply`

## Exercise 8.2
1. Create a list with the following elements - numeric vector, matrix, data frame.
2. Add a character vector to the above list without recreating the entire list.
3. Create a list of two lists. How would you access elements in this data structure?
