---
title: "Homework from Lab-09"
output: 
  html_document:
    df_print: paged
author: "my name here"
date: "`r Sys.time()`"
---

<style type="text/css">
.main-container {
  max-width: 1300px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r echo = FALSE}
######################################################################
##################### FOLLOW THESE DIRECTIONS ########################
### 1. Don't delete this code chunk, it won't appear in the output ###
### 2. Put your name in the author field above                     ###
### 3. Write the code for all questions in this R Notebook         ###
### 4. ALWAYS comment your code.                                   ###
### 5. Press the knit button when you're done, your code must      ###
###    run with no errors in order for this to work. If your       ###
###    code has errors, comment them out                           ###
######################################################################
######################################################################
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center", tidy = FALSE, max.print = 20)
```

---

Due date: Tuesday, Oct 12, 1:00 pm EST

```{r}
char_vec <- c("hello-1", "this-2", "is-3-1", "a-4-1", "char-5", "vector-6")
```


1. Extract the first three letters from each element of the character vector above. (2 pts)
2. Extract the last three letters from each element of the character vector above. (3 pts)
3. How would you create a *vector* of the following elements
  *A1, A2, A3, B1, B2, B3, C1, C2, C3, D1, D2, D3*? (2 pts)
4. How would you create the following *element* "A1+A2+A3+B1+B2+B3+C1+C2+C3+D1+D2+D3"? (3 pts)
