---
title: "Lab 19 - Statistics"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: TRUE
---

### Correlation
```{r, eval = F}
cor(x, y)
```

```{r, eval = F}
cor(x, y, method = c("pearson", "kendall", "spearman"))
```

- Using the $data.csv$ file, compute the correlation between RPF and mRNA.
- Use appropriate transformations.

### Regression
Simple linear regression
```{r, eval = F}
lm(y ~ x)
```

Multiple linear regression
```{r, eval = F}
lm(y ~ x1 + x2 + x3)
```

- Find out how mRNA abundances changes as a function of gene length.
- Find out how mRNA abundances changes as a function of various sequence features such as gene length, FEcap, uATG, utr, and gc.
- Use `tidy(), augment(), glance()`

### Random number generation

### Comparisons of mean
Student's t-test (equal variance)
```{r, eval = F}
t.test(x, mu = , alternative = c("two.sided", "less", "greater"), var.equal = TRUE) # One-sample
t.test(x, y, alternative = c("two.sided", "less", "greater"), var.equal = TRUE) # Two-sample
```

Welch t-test (unequal variance)
```{r, eval = F}
t.test(x, y, alternative = c("two.sided", "less", "greater"), var.equal = FALSE)
```

Paired t-test
```{r, eval = F}
t.test(x, y, alternative = c("two.sided", "less", "greater"), paired = TRUE)
```

Wilcoxon/Mann-Whitney signed-rank test
```{r, eval = F}
wilcox.test(x, y, alternative = c("two.sided", "less", "greater"))
```

### Comparisons of distributions
Kolmogorov-Smirnov test
```{r, eval = F}
ks.test(x, y, alternative = c("two.sided", "less", "greater"))
```

### Multiple hypothesis testing
- Bonferroni Correction
- False Discovery Rate (FDR)
 