---
title: "Lab 19 - Statistics"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: TRUE
---

### Correlation
```{r, eval = F}
cor(x, y)
```

```{r}
x <- 1:10
y <- sample(1:10)

x
y

cor(x, y)
cor(x, x)
```

```{r}
library(tidyverse)
```

```{r}
my_data <- data.frame(x = c(1, 2, 4, 10, 100, 101, 105, 200),
                      y = 1:8)
my_data

ggplot(my_data, aes(x = x, y = y)) +
  geom_point()
```

```{r}
cor(my_data$x, my_data$y, method = "pearson")
cor(my_data$x, my_data$y, method = "spearman")

```




```{r, eval = F}
cor(x, y, method = c("pearson", "kendall", "spearman"))
```

- Using the $data.csv$ file, compute the correlation between RPF and mRNA.
- Use appropriate transformations.

```{r}
yeast <- read_csv("data.csv")
yeast
```
```{r}
ggplot(yeast, aes(x = mRNA, y = RPF)) +
  geom_point()

ggplot(yeast, aes(x = mRNA, y = RPF)) +
  geom_point() + 
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```

```{r}
yeast %>%
  summarize(pearson.linear = cor(RPF, mRNA, method = "pearson"),
            pearson.log = cor(log(RPF), log(mRNA), method = "pearson"),
            spearman.linear = cor(RPF, mRNA, method = "spearman"),
            spearman.log = cor(log(RPF), log(mRNA), method = "spearman"))
```



### Regression
Simple linear regression
```{r, eval = F}
lm(y ~ x)
```

```{r}
cor(log(yeast$RPF), log(yeast$mRNA))
cor(log(yeast$mRNA), log(yeast$RPF))

rpf.on.mrna <- lm(log(RPF) ~ log(mRNA), data = yeast)
summary(rpf.on.mrna)
mrna.on.rpf <- lm(log(mRNA) ~ log(RPF), data = yeast)
summary(mrna.on.rpf)
```

```{r}
library(broom)
```

```{r}
rpf.on.mrna
class(rpf.on.mrna)

tidy(rpf.on.mrna)
glance(rpf.on.mrna)
augment(rpf.on.mrna)
```




Multiple linear regression
```{r, eval = F}
lm(y ~ x1 + x2 + x3)
```

```{r}
ggplot(yeast, aes(x = Length, y = mRNA)) +
  geom_point() +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  geom_smooth(method = "lm")
```

```{r}
mrna.vs.len <- lm(log(mRNA) ~ log(Length), data = yeast)
glance(mrna.vs.len)
tidy(mrna.vs.len)
```

```{r}
yeast
```

```{r}
full.mRNA.model <- lm(log(mRNA) ~ log(Length) + FEcap + uATG + log(utr) + gc, data = yeast)
glance(full.mRNA.model)
tidy(full.mRNA.model)


noUTR.mRNA.model <- lm(log(mRNA) ~ log(Length) + FEcap + uATG + gc, data = yeast)
glance(noUTR.mRNA.model)
tidy(noUTR.mRNA.model)
```

```{r}
ggplot(yeast, aes(x = gc)) +
  geom_density(fill = "black")
```

```{r}
library(MASS)
```

```{r}
stepAIC(full.mRNA.model)
```


- Find out how mRNA abundances changes as a function of gene length.
- Find out how mRNA abundances changes as a function of various sequence features such as gene length, FEcap, uATG, utr, and gc.
- Use `tidy(), augment(), glance()`

### Random number generation

```{r}
rand.uniform <- runif(10000)
hist(rand.uniform, breaks = 30)

rand.norm <- rnorm(10000)
hist(rand.norm)

x <- seq(-10, 10, 0.01)
norm.density <- dnorm(x, mean = 0, sd = 1)

plot(x, norm.density)
```




### Comparisons of mean
Student's t-test (equal variance)

```{r, eval = F}
t.test(x, mu = , alternative = c("two.sided", "less", "greater"), var.equal = TRUE) # One-sample
t.test(x, y, alternative = c("two.sided", "less", "greater"), var.equal = TRUE) # Two-sample
```

Welch t-test (unequal variance)
```{r, eval = F}
t.test(x, y, alternative = c("two.sided", "less", "greater"), var.equal = FALSE)
```

Paired t-test
```{r, eval = F}
t.test(x, y, alternative = c("two.sided", "less", "greater"), paired = TRUE)
```

Wilcoxon/Mann-Whitney signed-rank test
```{r, eval = F}
wilcox.test(x, y, alternative = c("two.sided", "less", "greater"))
```

Tests
```{r}
rand.normal.m10 <- rnorm(n = 100, mean = 10)
hist(rand.normal.m10)
mean(rand.normal.m10)
t.test(rand.normal.m10, mu = 10)
t.test(rand.normal.m10, mu = 11, alternative = "less")
t.test(rand.normal.m10, mu = 11, alternative = "greater")
```


### Comparisons of distributions
Kolmogorov-Smirnov test

```{r, eval = F}
ks.test(x, y, alternative = c("two.sided", "less", "greater"))
```



### Multiple hypothesis testing
- Bonferroni Correction
- False Discovery Rate (FDR)

```{r}
rand.pvalues <- rbeta(10000, shape1 = 0.4, shape2 = 0.4)
hist(rand.pvalues, breaks = 50)

sum(rand.pvalues < 0.05)

adj.pvalues.bonf <- p.adjust(rand.pvalues, method = "bonferroni")
sum(adj.pvalues.bonf < 0.05)

adj.pvalues.fdr <- p.adjust(rand.pvalues, method = "fdr")
sum(adj.pvalues.fdr < 0.05)
```

```{r}
rand.pvalues <- rbeta(10000, shape1 = 0.4, shape2 = 1)
hist(rand.pvalues, breaks = 50)
```




















 