---
title: "Lab 05 in-class"
output: 
  html_document:
    df_print: paged
---

# Tabulate and Summary
```{r}
# Read in the data frame
my_data <- read.table("~/homeworks/Homework-03/homework_data.tsv", header = TRUE, stringsAsFactors = FALSE)
my_data
```

```{r}
summary(my_data$Length) # Summarise one column of the data frame
summary(my_data) # Summarise all columns of the data frame
```


## Exercise 5.1
### 1. On which variable in our data frame will table be most useful?
```{r}
uATG_table <- table(my_data$uATG)
uATG_table

# table(my_data$TE) # Don't do this

uATG_table["TRUE"] # Elements in a vector can be referenced by their names
```
Only columns with a limited category of element types should be tabulated. 

### 2. What is the median RPF in our dataset?
```{r}
summary_RPF <- summary(my_data$RPF)
summary_RPF

summary(my_data$RPF)[3]

summary(my_data$RPF)["Median"] # Elements in a vector can be referenced by their names

summary_RPF[3]
summary_RPF["Median"] # Elements in a vector can be referenced by their names

names(summary_RPF) # Get names of a vector
```

### 3. What is the average mRNA value for genes with at most 400 codons?
```{r}
genes_below_400 <- my_data[my_data$Length <= 400, ] # Get all short genes <400
genes_below_400

summary_mrna_short <- summary(genes_below_400$mRNA) # Store the summary of mRNA into a vector
summary_mrna_short["Mean"]
```

## Exercise 5.2
### 1. What happens when you sample more values than the number of elements from a vector with replace=FALSE?
```{r}
vector_a <- 1:10
# sample(vector_a, size = 100, replace = FALSE)
```
Sample more values than the number of elements from a vector with replace=FALSE leads to an error.

### 2. Generate 1000 coin tosses from a fair coin and tabulate the results?
```{r}
coin <- c("Heads", "Tails")
coin
tosses <- sample(coin, size = 1000, replace = TRUE)
# tosses
table(tosses)
```

### 3. Generate 1000 coin tosses from a biased coin where the likelihood of HEADS is twice that of TAILS.
```{r}
biased_tosses <- sample(coin, size = 1000000, replace = TRUE, prob = c(2, 1))
table(biased_tosses)
```

### 4. Generate 1000 rolls of a biased die with the following odds
1	2	3	4	5	6
6	10	4	2	7	8
```{r}
die <- 1:6
prob_die <- c(6, 10, 4, 2, 7, 8)

rolls <- sample(die, size = 10000000, replace = TRUE, prob = prob_die)
roll_table <- table(rolls)
roll_table

# Check if the sampling process is consitent with the biases
roll_table/prob_die

```


## Exercise 5.3
Read in homework_data.tsv file from Lab 03
### 1. Check if 1246th gene in the dataset has more than 500 mRNA and is atleast 300 codons long.
```{r}
my_data

my_data[1246, ]

# Method 1
if(my_data[1246, "mRNA"] > 500 & my_data[1246, "Length"] >= 300){
  "Yes!"
}else{
  "Nope"
}

# Method 2
if(my_data$mRNA[1246] > 500 & my_data$Length[1246] >= 300){
  "Yes!"
}else{
  "Nope"
}


```

### 2. For all genes that are more than 400 codons long, print “LONG”, otherwise print “SHORT”. Tabulate your results.
```{r}
vector_longshort <- ifelse(my_data$Length > 400, "LONG", "SHORT")
table(vector_longshort)
```

### 3. How would you add/append this above result to the original dataset?
```{r}
# Method 1
new_data <- data.frame(my_data, 
                       size = vector_longshort)
new_data

# Method 2
my_data["size-2"] <- vector_longshort
my_data

# Method 3
my_data$size3 <- vector_longshort
my_data

# Method 4
cbind(my_data, vector_longshort)
```


## Exercise 5.4
### 1. Count the number of 4s in sample of 1000 unbiased die rolls using for loops.
```{r}
die <- 1:6
rolls <- sample(die, size = 1000, replace = TRUE)
# rolls

table(rolls)
sum(rolls == 4)

count4 <- 0               # Initialized a counter
for(dummy in rolls){      # Set up a for loop
  if(dummy == 4){         # Does the value of dummy equals 4
    count4 <- count4 + 1  # If it does, increase the counter by 1
  }
}
count4
```

### 2. Using for loops, create a new vector of differences between RPF and mRNA values of genes such that there are only positive values in this vector. Save this vector in the data frame.
```{r}
diff_values <- my_data$RPF - my_data$mRNA ## Don't use this

# Method 1 (Loop through one variable (RPF) and use a counter for the other variable (mRNA))
diff_vector <- NULL
diff_vector <- c()

counter <- 1
for(temp_rpf in my_data$RPF){
  diff <- temp_rpf - my_data$mRNA[counter]
  
  # If mRNA > RPF, change the sign and make the difference a positive value
  if(diff < 0){
    diff <- -diff
  }
  
  # Store the difference between RPF and mRNA into a difference vector
  diff_vector <- c(diff_vector, diff)
  
  counter <- counter + 1
}

diff_vector

# Method 2
new_diff_vec <- c()

for(rowPos in 1:nrow(my_data)){
  diff <- my_data$RPF[rowPos] - my_data$mRNA[rowPos]
  
  # If mRNA > RPF, change the sign and make the difference a positive value
  if(diff < 0){
    diff <- -diff
  }
  
  # Store the difference between RPF and mRNA into a difference vector
  new_diff_vec <- c(new_diff_vec, diff)
}

new_data <- data.frame(my_data, diff_vector, new_diff_vec)
new_data
```

### 3.
```{r}
diff_values <- my_data$RPF - my_data$mRNA 

# Method 1
all_pos_values <- ifelse(diff_values < 0, -diff_values, diff_values)
new_data <- data.frame(new_data, all_pos_values)
new_data

# Method 2 
abs_diff_values <- abs(my_data$RPF - my_data$mRNA)
new_data <- data.frame(new_data, abs_diff_values)
new_data
```

























