---
title: "Lab-03 Homework Solutions"
output: html_notebook
---

## Questions

Download the homework_data.tsv file and answer the following questions.
```{r}
# Load the home_data file
hw_data <- read.table("homework_data.tsv", header = TRUE, stringsAsFactors = FALSE)
hw_data
```


### 1. Create a new data frame with only genes longer than 400 codons and containing utrs.
```{r}
# Method 1
new_data <- hw_data[(hw_data$Length) > 400 & !is.na(hw_data$utr), ]
new_data

# Method 2
new_data1 <- hw_data[which(hw_data$Length > 400), ]
new_data2 <- new_data1[which(new_data1$utr > 0), ]
new_data2

# # Method 3
# new_data3 <- new_data1[which(new_data1$utr), ]
# new_data3
```

### 2. What is the difference in the average RPF values of genes with and without uATGs?
```{r}
# Method 1
genes_w_uATG <- hw_data[hw_data$uATG, ]
genes_w_uATG
genes_wo_uATG <- hw_data[!hw_data$uATG, ]
genes_wo_uATG

# Method 2
genes_w_uATG <- hw_data[hw_data$uATG == TRUE, ]
genes_w_uATG
genes_wo_uATG <- hw_data[hw_data$uATG == FALSE, ]
genes_wo_uATG

avg_RPF_w_uATG <- sum(genes_w_uATG$RPF)/nrow(genes_w_uATG)
avg_RPF_w_uATG
avg_RPF_wo_uATG <- sum(genes_wo_uATG$RPF)/nrow(genes_wo_uATG)
avg_RPF_wo_uATG

avg_RPF_w_uATG - avg_RPF_wo_uATG


avg_RPF_w_uATG <- mean(genes_w_uATG$RPF)
avg_RPF_w_uATG
avg_RPF_wo_uATG <- mean(genes_wo_uATG$RPF)
avg_RPF_wo_uATG

avg_RPF_w_uATG - avg_RPF_wo_uATG

```

### 3. What is the length of a genome which has only the genes with mRNA abundances greater than 250?
```{r}
mrna_gt_250 <- hw_data[hw_data$mRNA > 250, ]
mrna_gt_250

sum(mrna_gt_250$Length)

# Method 2
pos_mrna_gt_250 <- which(hw_data$mRNA > 250)

sum(hw_data[pos_mrna_gt_250, "Length"])
sum(hw_data$Length[pos_mrna_gt_250])
```

### 4. What proportion of total mRNA is due to all short genes (less than 250 codons)?
```{r}
short_genes <- hw_data[hw_data$Length < 250, ]
long_genes <- hw_data[hw_data$Length >= 250, ]
short_genes
long_genes

mrna_short <- sum(short_genes$mRNA)

mrna_short/sum(hw_data$mRNA)
```

### 5. What is the average GC content of utr containing short genes (less than 250) ?
```{r}
short_genes <- hw_data[hw_data$Length < 250, ]
short_w_utr <- short_genes[!is.na(short_genes$utr), ]

mean(short_w_utr$gc)

# 1
short_w_utr_noNAgc <- short_w_utr[!is.na(short_w_utr$gc), ] # Remove any rows that have NA in gc column
mean(short_w_utr_noNAgc$gc)

# 2
mean(short_w_utr$gc, na.rm = TRUE)

```


















