---
title: "Quiz 3 answers"
output: html_document
---

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
pokemon <- read_csv("pokemon_go_captures.csv")
pokemon
```


### 1. Find out which height_bin has the smallest combat_power on average.
```{r}
pokemon %>%
  group_by(height_bin) %>%
  summarise(mean.combat = mean(combat_power)) %>%
  arrange(mean.combat) %>%
  .[1,]
```

### 2. Assume that the total awesomeness of a pokemon is the product of combat_power and hit_points. Which pokemon species has the highest awesomeness?
```{r}
pokemon %>%
  group_by(species) %>%
  mutate(awesomeness = combat_power*hit_points) %>%
  arrange(desc(awesomeness)) 

pokemon %>%
  mutate(awesomeness = combat_power*hit_points) %>%
  arrange(desc(awesomeness)) %>%
  select(species, awesomeness, everything())
```

### 3. Plot height vs weight of all pokemons. Make appropriate transformations of axes and points. Add a correlation coefficient and regression line to the plot.
```{r}
ggplot(data = pokemon, aes(x = weight_kg, y = height_m)) +
  geom_point(alpha = 0.3) +
  labs(x = "Weight (Kg)", y = "Height (m)") +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  geom_smooth(method = "lm", col = "red") +
  stat_cor(aes(label = ..rr.label..))
```


### 4. Make a boxplot of combat_power (y) for weight_bin (x) category. Compare the means of combat_power betwen “extra_large” and “extra_small” weight_bin and plot the p-value on the boxplot.
```{r}
ggplot(data = pokemon, aes(x = weight_bin, y = combat_power)) +
  geom_boxplot() +
  labs(x = "Weight category", y = "Combat power") +
  stat_compare_means(comparisons = list(c("extra_large", "extra_small")), method = "t.test")

ggplot(data = pokemon, aes(x = weight_bin, y = combat_power)) +
  geom_boxplot() +
  scale_y_continuous(trans = "log10") +
  labs(x = "Weight category", y = "Combat power") +
  stat_compare_means(comparisons = list(c("extra_large", "extra_small")), method = "t.test")
```

### 5. Some pokemons have a unique fast_attack (not shared by any other pokemon). How many pokemons have a unique fast_attack? Bonus (+1) Which pokemons have a unique fast_attack?
```{r}
# Method 1
pokemon %>%
  group_by(fast_attack) %>%
  tally() %>%
  filter(n == 1)

unq.fast <- pokemon %>%
  group_by(fast_attack) %>%
  tally() %>%
  filter(n == 1) %>%
  pull(fast_attack)
unq.fast

# Method 2
pokemon %>%
  filter(fast_attack %in% unq.fast)

unq.fast.df <- pokemon %>%
  group_by(fast_attack) %>%
  tally() %>%
  filter(n == 1)
unq.fast.df

left_join(unq.fast.df, pokemon, by = "fast_attack")

# Method 3
pokemon %>%
  group_by(fast_attack) %>%
  tally() %>%
  filter(n == 1) %>%
  left_join(., pokemon, by = "fast_attack")
```















