---
title: "Lab 01 - in-class exercises"
author: "Premal Shah"
output: html_notebook
---

This is where the next piece of code will be written.
```{r}
2+2
```

## R is a giant calculator
```{r}
2+2       # Simple addition
4*5*(7+9) # Complex calculation
sqrt(96)  # Square root of 96
exp(4)    # Exponentials e^4
log(10)   # log of 10 (base e)
log(10, base=10) # log of 10 (base 10)
```
When you run code chunks inside the notebook, all the output ends up together. However, in the *knitted* html file, each output comes right after the command.

## Variables in R
```{r}
variable_x <- 4 # Assign a value of 4 to variable 'x'
variable_y <- 5 # Assign a value of 4 to variable 'y'
```
No output is returned during variable assignment.

```{r}
variable_x # To see the value of a variable just type its name
```

### Working directly with variables
```{r}
variable_x + variable_y
```

```{r}
variable_z <- 10
```

```{r}
variable_z
```

### Creating new variables to store values from arithmetic operations performed on other variables
```{r}
variable_z <- variable_x + variable_y 
variable_z
```

Changing the value of variable_x will not change the value of variable_z, until you rerun the arithmetic operation. Once a value is saved in a variable, it has no memory of how the value was arrived at. There is no auto-updating of variables.

```{r}
variable_x <- 17 # New value for variable_x
variable_z # Value of variable_z remains unchagned

variable_z <- variable_x + variable_y # Only after this operation, the value of variable_z will change
variable_z
```

## Logical statements - comparing variables
```{r}
variable_x == variable_y # Are these two variables equal
variable_x <- 1
variable_y <- 0.99999999999999
variable_x == variable_y # Are these two variables equal
```
The == sign requires identical output to return a `r TRUE` statement. It does not approximate values.

```{r}
variable_x <- 1
variable_y <- 0.999999
```

```{r}
variable_x != variable_y # Check if variable_x is NOT EQUAL to variable_y
```

```{r}
variable_x =! variable_y
variable_x
```
The order of operation is important. If  you switch != to =!, instead of checking for inequality, you are assigning a new value to variable_x. In this case, the value of NOT (!) which is stored as a logical FALSE value.


```{r}
## Two ways to assign a value to a variable
var_a <- 4
var_b = 4
```

```{r}
variable_y
!variable_y
```

```{r}
variable_x <- 5
variable_x <= variable_y # Is x less than or equal to y
variable_x < variable_y # Is x less than y
```

### Complex comparisons between sets of variables
```{r}
(variable_x == variable_y) | (variable_x > variable_y) # Is x equal to y OR x greater than y
```

```{r}
(variable_x != variable_y) & (variable_x < variable_y) # Is x different from y AND x less than y
```

## Exercise 1.1
### 1. Create three variables and check if either of the first two variables are greater than the third.
```{r}
# Three new variables
var_a <- 10
var_b <- 20
var_c <- 15
```

```{r}
(var_a > var_c) | (var_b > var_c) # Check if A greater than C OR if B greater C -- This is correct
(var_a | var_b) > var_c # Check if (A OR B) is greater than C -- This is incorrect
```

When asking var_a OR var_b, you are ONLY checking if the two varaibles exist, in which case the program will return a TRUE statement if the two variables have been previously defined.
```{r}
var_a | var_b 
4 | 5
```

### 2. Create three variables and check if the sum of the first two variables is greater than the third variable.
```{r}
# Both the following codes will give the right answer. 
# However, always use parentheses for clarity of code and to minimize errors
(var_a + var_b) > var_c # Right
var_a + var_b > var_c # Wrong (but works)
```

### 3. Create four variables and check if the product of the first two variables is the same as the product of the last two variables.
```{r}
var_d <- 7

(var_a * var_b) == (var_c * var_d)
```

## Types of variables
```{r}
variable_u <- TRUE # Logical
variable_v <- 13 # Numeric
variable_w <- 10L # Integer
variable_x <- 3+4i # Complex numbers (we will ignore these for this course)
```

### Comparing strings
```{r}
string_1 <- "hey"
string_2 <- "hey"

string_1 == string_2

string_1 <- "hey"
string_2 <- "they"

string_1 == string_2

```
Logical statements can be used to compare strings.

## Exercise 1.2
### 1. What results when you add TRUE variable with a numeric?
```{r}
var_a <- 10 # Numeric varaible
var_b <- TRUE # Logical variable

var_a + var_b
```
Since the numeric value of a TRUE variable is 1, adding it to an integer or a numeric variable will increase the variable's value by 1.

### 2. What results when you multiply FALSE with an integer?
```{r}
var_a <- 10L # Integer variable
var_a * FALSE 
TRUE + TRUE
```
Since the numeric value of a FALSE variable is 0, multiplying it to any integer or a numeric variable will return a 0.

### 3. What results when you add NA to numeric?
```{r}
var_a + NA 
```
NA is a special case of logical variables. Any arithmetic operations with it will result in an NA.


### 4. What is the difference in output (if any) between adding a numeric to a complex number and adding an integer to a complex number?
```{r}
var_numeric <- 10 # Numeric 
var_integer <- 10L # Integer
var_complex <- 3+5i # Complex
```

```{r}
var_numeric + var_complex
var_integer + var_complex
```
Adding a numeric or an integer value to a complex will change the Real part of the complex number without affecting the Imaginary part.

### 5. What is the difference when you add a numeric to a complex number versus multiplying a numeric to a complex number? 
```{r}
var_numeric * var_complex
```
Multiplying a numeric or an integer value to a complex will change both the Real and the Imaginary parts, consistent with complex number arithmetic.






