---
title: "Quiz 01"
output:
  html_document: default
  html_notebook: default
---

- Each question is 1 point.
- Write the `codes` for all questions in a single R Notebook (file).
- ALWAYS comment your code.
- Name the file as "Lastname_Lab04_quiz.nb.html"

1. Create two numeric vectors of length 3. Create a new combined vector EXCLUDING the second elements of both vectors.
2. Create a vector of all multiples of 7 between 14 and 91 in DECREASING order?
3. Create any 3x3 numeric matrix. Create a vector of sums of elements in each row using two different methods.
4. Why do you get an error when you add two variables WITHOUT assigning them values?
5. & 6. Create the following vectors using `rep` function
```{r, echo=FALSE}
rep(14:17,each=3)
rep(sample(1:4),3)
# library(knitr)
# q <- as.data.frame(t(rep(" ",12)))
# kable(q,col.names=rep(14:17,each=3))
# kable(q,col.names=rep(sample(1:4),3))
```

7. Create a new data frame shown below
```{r, echo=FALSE}
# Creating data frames
my_data <- data.frame(
    Gene = c("Gene1", "Gene2", "Gene3"),
    Length = c(100L,200L,300L),
    Expression = floor(rnorm(3)*100),
    Disease = c(TRUE, FALSE, FALSE),
    Strand = c("+","+","-"),
    stringsAsFactors = FALSE
)
my_data
```

8. What does the function `dim()` do?

Read in the quiz_data.tsv file and answer the following

9. What is the size of the genome comprising of only the last 100 genes in the dataset?

10. Where in the dataset are the following genes - YBL006C, YBL025W, and YAL025C?
