---
title: "Shah Lab04 quiz"
output: html_notebook
---

### 1. Create two numeric vectors of length 3. Create a new combined vector EXCLUDING the second elements of both vectors.
```{r}
vec_1 <- c("A", "B", "C")
vec_2 <- c("L", "M", "N")

comb_vector <- c(vec_1[-2], vec_2[-2])
comb_vector

comb_vector1 <- c(vec_1[1], vec_1[3], vec_2[1], vec_2[3])
comb_vector1
```


### 2. Create a vector of all multiples of 7 between 14 and 91 in DECREASING order?
```{r}
seq(from = 91, to = 14, by = -7)
```

### 3. Create any 3x3 numeric matrix. Create a vector of sums of elements in each row using two different methods.
```{r}
num_matrix <- matrix(1:9, nrow = 3)
num_matrix

# Method 1
sum_vector1 <- c(sum(num_matrix[1, ]), sum(num_matrix[2, ]), sum(num_matrix[3, ]))
sum_vector1

# Method 2
sum_vector2 <- rowSums(num_matrix)
sum_vector2

# Method 3
sum_vector3 <- num_matrix[, 1] + num_matrix[, 2] + num_matrix[, 3]
sum_vector3
```


### 4. Why do you get an error when you add two variables WITHOUT assigning them values?
```{r}
var_a + var_b
```

### 5. [1] 14 14 14 15 15 15 16 16 16 17 17 17
```{r}
rep(x = 14:17, each = 3)
```

### 6. [1] 2 3 4 1 2 3 4 1 2 3 4 1
```{r}
new_vec <- c(2, 3, 4, 1)
rep(x = new_vec, times = 3)
```

### 7. Create a new data frame shown below
##    Gene Length Expression Disease Strand
## 1 Gene1    100        -62    TRUE      +
## 2 Gene2    200        -24   FALSE      +
## 3 Gene3    300         23   FALSE      -
```{r}
my_df <- data.frame(
  Gene = c("Gene1", "Gene2", "Gene3"),
  Length = c(100L, 200L, 300L),
  Expression = c(-62, -24, 23),
  Disease = c(TRUE, FALSE, FALSE),
  Strand = c("+", "+", "-"),
  stringsAsFactors = FALSE
)
my_df
```



### 8.What does the function dim() do?
```{r}
?dim
help(dim)
```

### 9. What is the size of the genome comprising of only the last 100 genes in the dataset?
```{r}
quiz_data <- read.table("quiz_data.tsv", header = TRUE, stringsAsFactors = FALSE)
quiz_data

# Method 1
last_100_1 <- (length(quiz_data$Length)-99):length(quiz_data$Length)

sum(quiz_data$Length[last_100_1])

# Method 2
length(4739:4839)
sum(quiz_data$Length[4739:4839])
sum(quiz_data$Length[4740:4839])

# Method 3
sum(tail(quiz_data$Length, n = 100))

tail(quiz_data)

# Method 4
sum(quiz_data$Length[-(1:4739)])
```

### 10. Where in the dataset are the following genes - YBL006C, YBL025W, and YAL025C?
```{r}
genes_of_int <- c("YBL006C", "YBL025W", "YAL025C")
which(quiz_data$Gene %in% genes_of_int)
```

















