---
title: "Lab-11 in-class"
output: html_document
---

```{r}
library(IRanges)
```

```{r}
var.named.irange <- IRanges(start = 4:6, 
                            width = 4,
                            names = c("Gene1", "Gene2", "Gene3"))
var.named.irange
```
```{r}
var.named.irange[1]
var.named.irange["Gene3"]
```

## Exercise 11.1
### 1. Imagine a genome consisting of only five consecutive genes of lengths 100, 200, 300, 400 and 500, in that order. How would you represent this genome as an IRanges object?
```{r}

gene.widths <- c(100, 200, 300, 400, 500)
vec.start <- 1

for(dummy in 1:(length(gene.widths)-1)){
  vec.start[dummy + 1] <- vec.start[dummy] + gene.widths[dummy]
}

vec.start
gene5.irange <- IRanges(start = vec.start, 
        width = gene.widths, 
        names = c("G1", "G2", "G3", "G4", "G5"))
gene5.irange
```


### 2. Create an IRanges genome using the genes and their lengths in yeast_data.csv file.
```{r}
yeast_data <- read.table("yeast_data.tsv", header = TRUE)
yeast_data

yeast.widths <- yeast_data$Length

vec.start <- 1

for(dummy in 1:(length(yeast.widths)-1)){
  vec.start[dummy + 1] <- vec.start[dummy] + yeast.widths[dummy]
}

IRanges(start = vec.start, 
        width = yeast.widths, 
        names = yeast_data$Gene)
```

## Exercise 11.2

### 1. Starting with the 5-gene genome from the previous exercise, extend the lengths of first 3 genes at the 5' ends and last 2 genes at their 3' ends by 50bp.
```{r}
gene5.irange

new.gene5.irange <- resize(gene5.irange,
                           width = width(gene5.irange) + 50,
                           fix = c(rep("end", 3), rep("start", 2)))
new.gene5.irange

new.gene5.irange <- resize(gene5.irange,
                           width = width(gene5.irange) + 50,
                           fix = c("end", "end", "end", "start", "start"))
new.gene5.irange
```

### 2. Assuming all 5 genes have a promoter of 75 bp. Create a new IRanges object for the promoters.
```{r}
promoter5.irange <- flank(gene5.irange, 
      width = 75)
names(promoter5.irange) <- c("pG1", "pG2", "pG3", "pG4", "pG5")

c(promoter5.irange, gene5.irange)

```




















