---
title: "Lab-14 Introduction to Bioconductor - Rle and IRanges"
author: "Premal Shah"
date: "Oct 20, 2020"
output: 
  html_notebook:
    theme: spacelab
    toc: yes
    toc_float: yes
---

## Bioconductor 
[Website](https://www.bioconductor.org/)

## Installing Bioconductor packages (in console)

```{r, eval=F}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```

## S4 objects

## Running length encoding

## IRanges

### Create an IRanges object
### Structure (S4) of an IRanges object
### Working with IRanges
### Combining IRanges

## Exercise 14.1
1. Imagine a genome consisting of only five consecutive genes of lengths 100, 200, 300, 400 and 500, in that order. How would you represent this genome as an IRanges object?

## Manipulating IRanges

```{r, eval=F}
shift(IR, shift =)

flank(IR, width =, start = T/F)

resize(IR, width =, fix = "start"/"end"/"center")

gaps()
```

## Exercise 14.2
1. Starting with the 5-gene genome from the previous exercise, extend the lengths of first 3 genes at the 5' ends and last 2 genes at their 3' ends by 50bp.
2. Assuming all 5 genes have a promoter of 75 bp. Create a new IRanges object for the promoters.
