---
title: "Lab 06 in-class"
output: html_document
---

## Exercise 6.1
### 1. Keep rolling an unbiased die until you see a 6. How many dice rolls did it take?
```{r}
# First - throw a die
roll <- sample(1:6, size = 1)
# Keep track of how many rolls
counter <- 1
print(roll)

while(roll != 6){ # Check if the current throw is a 6
  roll <- sample(1:6, size = 1) # If not 6, then throw another roll
  counter <- counter + 1 # Increase the number of rolls by 1
  print(roll)
}
counter
```

### 2. Repeat the same process for a biased dice with the following odds.
1	2	3	4	5	6
10	9	8	7	6	5
```{r}
odds <- c(10, 9, 8, 7, 6, 5)
# First - throw a die
roll <- sample(1:6, size = 1, prob = seq(from = 10, to = 5, by = -1))
# Keep track of how many rolls
counter <- 1
print(roll)

while(roll != 6){ # Check if the current throw is a 6
  roll <- sample(1:6, size = 1, prob = odds) # If not 6, then throw another roll
  counter <- counter + 1 # Increase the number of rolls by 1
  print(roll)
}
counter
```

### 3. Repeat the above process until you see a 6 or you have rolled the dice 10 times, whichever comes first.
```{r}
odds <- c(10, 9, 8, 7, 6, 5)
# First - throw a die
roll <- sample(1:6, size = 1, prob = seq(from = 10, to = 5, by = -1))
# Keep track of how many rolls
counter <- 1
print(roll)

while(roll != 6 & counter < 10){ # Check if the current throw is a 6
  roll <- sample(1:6, size = 1, prob = odds) # If not 6, then throw another roll
  counter <- counter + 1 # Increase the number of rolls by 1
  print(roll)
}
counter
```


## Exercise 6.2
### 1. Write a function to calculate factorial of a value.
```{r}
# Method 1
my_factorial1 <- function(n){
  fact <- 1 # Initialize variable to store the value of the factorial
  
  # This starts at the values n and then decreases the value of n with each interation
  # For instance for n = 5, it starts at 5, then goes to 4, 3, 2, and 1
  while(n >= 1){ 
    fact <- fact * n # Update the value of the factorial
    n <- n - 1 # Decrease n by 1
  }
  return(fact)
}

# Method 2
my_factorial2 <- function(n){
  fact <- 1 # Initialize variable to store the value of the factorial
  
  # This creates a dummy variable whose value starts at 1 and then increases with each interation
  for(dummy in 1:n){ # Either 1:n OR seq(n) OR seq(from = 1, to = n)
    fact <- fact * dummy
  }
  
  return(fact)
}

# Method 3
my_factorial3 <- function(n){
  temp_vec <- 1:n # This creates a temporary vector of all numbers between 1 and n
  fact <- prod(temp_vec) # Take the product of all the elements
  
  return(fact)
}

```

```{r}
# Check the output
factorial(5) # In-built factorial function
my_factorial1(n = 5)
my_factorial2(n = 5)
my_factorial3(n = 5)
```


### 2. Write a function to calculate the fibonacci series up to n elements.
```{r}
# Method 1
fib_series <- function(n){
  fib_vec <- c(0, 1) # The first two elements of the series are fixed.
  
  # If the user requests first or first two elements, return the values 
  # directly from the initial vector
  if(n < 3){ 
    return(fib_vec[1:n])
  }else{ # If the use requests more elements
    y <- 3
    while(y <= n){
      fib_vec <- c(fib_vec, fib_vec[y-1]+fib_vec[y-2]) # Add the sum of the previous two elements to the existing vector. This grows the fib_vec by one more element
      y <- y + 1 # Increase the counter
    }
  }
  
  return(fib_vec)
}

fib_series2 <- function(n){
  fib_vec <- c(0, 1) # The first two elements of the series are fixed.
  
  # If the user requests first or first two elements, return the values 
  # directly from the initial vector
  if(n < 3){
    return(fib_vec[1:n])
  }else{
    for(dummy in 3:n)
      fib_vec <- c(fib_vec, fib_vec[dummy-1] + fib_vec[dummy-2]) # Add the sum of the previous two elements to the existing vector. This grows the fib_vec by one more element
    }
  return(fib_vec)
}
```

```{r}
fib_series(n = 10)
fib_series2(n = 10)
```

## Lists
```{r}
my_list <- list(
    a = 1:5,
    b = c("Gene1", "Gene2"),
    mat = matrix(1:4, nrow = 2)
)
my_list
```

```{r}
# Access elements like a dataframe
my_list$b
```

```{r}
# Access elements like a dataframe
my_list$mat
```

```{r}
my_list$mat[1, ]
```

## Exercise 6.3
### 1. Create a list with the following elements - numeric vector, matrix, data frame.
```{r}
new_list <- list(
  num_vec = c(1, 4, 5),
  num_matrix = matrix(1:4, nrow = 2, byrow = TRUE),
  my_data = data.frame(
    Gene = c("gene1", "gene2"),
    Length = c(100, 500)
  )
)

new_list
```

### 2. Add a character vector to the above list without recreating the entire list.
```{r}
# Method 1
new_list$char_vec <- c("A", "B", "D", "G", "K")
new_list

# Method 2
char_vec <- c("A", "B", "D", "G", "K")
new_list[[5]] <- char_vec
new_list

# Method 3
new_list[["new_char_vec"]] <- char_vec
new_list
```
Methods 1 and 3 create named elements. Method 2 creates an unnamed element.

### Difference between using single brackets and double brackets
```{r}
# When you use a single bracket, you get the output in the form of a list
new_list[1]
class(new_list[1])

# When you use a double bracket, you get the output in the dataclass of the element
new_list[[1]]
class(new_list[[1]])
```

```{r}
new_list$num_matrix[2, 2]
new_list[["num_matrix"]][2, 2]
# new_list["num_matrix"][2, 2] - This creates an error because the output is a list

```


### 3. Create a list of two lists. How would you access elements in this data structure?
```{r}
new_list2 <- list(
  num_vec = c(1, 4, 5),
  num_matrix = matrix(1:4, nrow = 2, byrow = TRUE)
)
new_list2

small_list <- list(
  a = 1:2,
  b = c("A", "B", "C")
)
small_list
```

```{r}
# Use the list() function to combine two lists into a list of list
list_of_list <- list(new_list2, small_list)
list_of_list
```


```{r}
# Accessing elements in a list of list
list_of_list[[1]][[2]][2, 2]
list_of_list[[1]]$num_matrix[2, 2]

list_of_list[[2]]$b[3]
list_of_list[[2]][[2]][3]
```
























