---
title: "Lab-02 in-class"
author: "Premal Shah"
output: html_notebook
---


## Vectors
```{r}
numeric_vector <- c(1, 2, 3, 4) # Vector of numbers
char_vector <- c("hello!", "this", "is", "a", "character", "vector")

numeric_vector
char_vector
```

## Exercise 2.1
### 1. Find the sum of first and last element of a vector.
```{r}
vector_numeric <- c(11, 12, 13, 14)

# Method 1
vector_numeric[1] + vector_numeric[4] # Access the first and last values by referencing the locations directly

# Method 2
vec_length <- length(vector_numeric) # When you don't know the location of the last element, use length() to find out the length of a vector
vec_length
vector_numeric[1] + vector_numeric[vec_length] # Use the new variable vec_length to reference the last element
vector_numeric[1] + vector_numeric[length(vector_numeric)] # This can also be done without saving the length of a vector in a new variable

# Method 3
sum(vector_numeric[1], vector_numeric[vec_length]) 
```

### 2. Create two numeric vectors of length 3 and combine them into a third vector.
```{r}
vector_1 <- c(1, 4, 7)
vector_2 <- c(11, 14, 17)

c(vector_1, vector_2) # c() functions allows you to combine multiple elements, and vectors.
```

### 3. What happens when you combine a character vector with a numeric vector?
```{r}
numeric_vector
char_vector

comb_vector <- c(numeric_vector, char_vector)
comb_vector 

numeric_vector[1] + numeric_vector[2]
comb_vector[1] + comb_vector[2]
```

When you combine numeric and character vectors, the numbers are stored as character elements now. You can no longer perform arithmetic operations on these elements.

### 4. Create two numeric vectors of length 3. Create a new vector with ONLY the second elements of both vectors.
```{r}
vector_1
vector_2

new_vec <- c(vector_1[2], vector_2[2])
new_vec
```

### 5. What happens when you reference an index outside the length of a vector?
```{r}
vector_1
vector_1[5]
```
When you reference an index outside the length of a vector, you get an NA value.

### 6. What happens when you reference a negative index?
```{r}
vector_1[-1]
vector_1[-5]
vector_1[0]
```
Negative indices allows you to print elements of a vector with specific elements removed. Whatever value is stored in location of the positive value of the index gets removed. For instance -2 index will remove the element at position 2, -3 will remove the element at position 3, and so on and so forth. 

### 6* How do you access or remove multiple elements of a vector?
```{r}
char_vector
char_vector[c(1, 3)]
char_vector[c(-2, -4)]
```

### 7. What happens when you add two numeric vectors of the same length?
```{r}
vector_1
vector_2

vector_1 + vector_2
```
When you add two vectors of the same length, the elements in the corresponding positons get added up in the new vector.

### 8. What happens when you add two numeric vectors of the different lengths?
```{r}
vector_1
numeric_vector <- c(11, 12, 13, 14, 15, 16, 17)
numeric_vector
vector_1 + numeric_vector
```
When you add two vectors of different length, the smaller vector gets recycled.

## Exercise 2.2
### 1. Create a vector of numbers from 1 to 100.
```{r}
# Method 1
vec_100_1 <- seq(from = 1, to = 100, by = 1)
vec_100_1

# Method 2 
vec_100_2 <- seq(from = 1, to = 100)
vec_100_2

# Method 3
vec_100_3 <- seq(to = 100)
vec_100_3

# Method 4
1:100
```

### 2. Create a vector of only even numbers between 1 and 100.
```{r}
vec_even <- seq(from = 2, to = 100, by = 2)
vec_even
```

### 3. From the vector of all numbers (1 to 100), create a new vector with only multiples of 5.
```{r}
seq(5, 100, 5) # Wrong - While this will give the right answer but the question was about subsetting a previous vector.
vec_100_1[seq(from = 5, to = 100, by = 5)] # Correct answer

```

### 4. Create a vector of all multiples of 10 between 1 and 100 in DECREASING order?
```{r}
vec_dec_10 <- seq(from = 100, to = 1, by = -10)
vec_dec_10
seq(from = 100, to = 0, by = -10) # If you stop at 0, you will get an additional element in the new vector
```

### 5. What values are in the 3rd and 7th element of the above vector?
```{r}
vec_dec_10[c(3, 7)]
```

### 6. What happens when you seq(10)?
```{r}
seq(10)
```

## Exercise 2.3
### 1. 
```{r}
rep(x = 5, times = 10)
```

### 2. 
```{r}
vec_5 <- 1:5
rep(x = vec_5, times = 2)
rep(x = seq(5), times = 2)
rep(x = 1:5, times = 2)
```

### 3.
```{r}
c(rep(x = 1, times = 2), rep(x = 2, times = 2), rep(x = 3, times = 2), rep(x = 4, times = 2), rep(x = 5, times = 2)) # While this works, its very cumbersome to write for longer vectors
rep(x = 1:5, each = 2) # Use each argument instead of times argument
```

### 4.
```{r}
c(rep(x = 1, times = 4), rep(x = 2, times = 3), rep(x = 3, times = 2), rep(x = 4, times = 1)) # While this works, its very cumbersome to write for longer vectors 

rep(x = 1:4, times = 4:1) # times argument can take a vector
rep(x = seq(from = 1, to = 4, by = 1), 
    times = seq(from = 4, to = 1, by = -1)) # A more explicit way of writing the answer.

```

## Exercise 2.4
### 1.
```{r}
new_matrix <- matrix(data = 1:15, nrow = 3, ncol = 5, byrow = TRUE)
new_matrix
```

### 2. Save the 4th column as a new vector
```{r}
new_vec <- new_matrix[, 4]
new_vec
```

### 3. Create a new vector where each element is the sum of 2nd and 3rd row.
```{r}
sum(new_matrix[2, ], new_matrix[3, ]) # WRONG. Using the sum function combines all the elements of the two rows into a single vector and sums it up.
new_matrix[2, ] + new_matrix[3, ] # Right
```


### 4. Subset a new 2x2 matrix from the above matrix with the following values
```{r}
sub_matrix <- new_matrix[c(2, 3), c(2, 4)]
sub_matrix
```











