---
title: "Lab-02 Homework"
output: html_notebook
---

## 1. Create a 3x3 numeric matrix. Rearrange the elements of this matrix using a single statement, such that no element is in the same place as the original matrix.
```{r}
new_matrix <- matrix(LETTERS[1:9], nrow = 3) # Create a 3x3 matrix
new_matrix 

new_matrix[c(3, 1, 2), ] # Shuffling either rows or columns will work
new_matrix[c(3, 1, 2), 1]
```

## 2. Create a 3x4 numeric matrix and then transpose it to a new matrix. https://en.wikipedia.org/wiki/Transpose
```{r}
numeric_matrix <- matrix(11:22, nrow = 3, ncol = 4) # Create a 3x4 matrix
numeric_matrix

t(numeric_matrix) # While the in-built transpose function works, I was look for answers based on functions we worked on in class.

matrix(numeric_matrix, nrow = 4, byrow = TRUE) # Refactoring a matrix by changing byrow status transposes the matrix.
```

## 3. Create a 3x4 numeric matrix and then store the sums of its rows into a vector. Store the sums of its columns into another vector.

```{r}
numeric_matrix

# Method 1 - Sum each row separately, store the values in variables and create a new vector
row1_sum <- sum(numeric_matrix[1, ])
row2_sum <- sum(numeric_matrix[2, ])
row3_sum <- sum(numeric_matrix[3, ])

sums_of_rows <- c(row1_sum, row2_sum, row3_sum)
sums_of_rows


# Method 2 - Create a new vector by summing each row separately 
sums_of_rows2 <- c(sum(numeric_matrix[1, ]), sum(numeric_matrix[2, ]), sum(numeric_matrix[3, ]))
sums_of_rows2


# Method 3 - Extract each column from the matrix as a vector and sum the four vectors
sums_of_rows3 <- numeric_matrix[, 1] + numeric_matrix[, 2] + numeric_matrix[, 3] + numeric_matrix[, 4]
sums_of_rows3

# Method 4 - in-built function for summing rows (didn't expect you to know this beforehand)
rowSums(numeric_matrix)
```

## 4. Perform standard arithmetic operations (addition, subtraction, multiplication and division) on two 2x2 matrices and store the values in new matrices and describe them. What happens when the matrices are of different sizes (2x2 and 3x3)?
```{r}
first_mat2x2 <- matrix(11:14, nrow = 2)
second_mat2x2 <- matrix(1:4, nrow = 2)

first_mat2x2
second_mat2x2

## Addition
first_mat2x2 + second_mat2x2

## Subtraction
first_mat2x2 - second_mat2x2

## Multiplication
first_mat2x2 * second_mat2x2

## Division
first_mat2x2 / second_mat2x2

mat_3x3 <- matrix(1:9, nrow = 3)

first_mat2x2 + mat_3x3
```
Matrices of different sizes (ranks) are non-conformable.

## 5. Rearrage the following matrix such that all the rows sum to the same value. (Not graded)
```{r}
new_matrix <- matrix(1:9, nrow = 3)
new_matrix

# Install magic package once
# install.packages("magic")

library(magic)# Load the magic package in the environment

magic(3)
magic(6)

rowSums(magic(3))
colSums(magic(3))
```










