---
title: "Lab 05 - Data frames and conditional statements"
author: "Premal Shah"
date: "Sep 23, 2021"
output: 
  html_notebook:
    theme: spacelab
    toc: yes
    toc_float: yes
---


## Dataframes
- What if you wanted different _types_ of data in each column?

### Creating a data frame
```{r}
my_data_frame <- data.frame(Gene = c("gene1", "gene2", "gene3"),
                            Length = c(100, 200, 300),
                            Disease = c(TRUE, FALSE, TRUE))
my_data_frame
```


### Accessing and subsetting data frames
- Like a matrix
- Using column names

## Reading external files

### Working with vectors and data frames
```{r, eval=FALSE}
# Length gives you number of columns
length()

# Dimensions of a data frame
dim()

# Number of rows of a data frame
nrow()

# Number of columns of a data frame
ncol()

# Identify type of objects
class()

# Structure of a data frame
str()

# Summary functions
min()
max()
mean()
sd()

# Summarize data in all columns
summary()

# Check the type of data in a column
is.logical()
is.numeric()
is.na()
```

## Exercise 5.1
1. Create a new data frame which contains only the first three columns of the old data frame.
2. How many observations have __daily_vaccinations__ values in them?

## Vectorized operations (detour)

## Locating and extracting values
- Using logical variables
- `which` function
- `%in%` function

## Exercise 5.2
1. How many observations are there for the `New Jersey` in the data frame?
2. Which state has the highest number of __people_vaccinated__? 
3. Is the state with the most number of __people_vaccinated__ also the state with the most __people_fully_vaccinated_per_hundred__?

## Tabulate

## Sample
- Sampling with/without replacement
- Biased sampling

### Exercise 5.3
1. What happens when you sample more values than the number of elements from a vector with `replace=FALSE`?
2. Generate 1000 coin tosses from a fair coin and tabulate the results?
3. Generate 1000 coin tosses from a biased coin where the likelihood of HEADS is twice that of TAILS.
4. Generate 1000 rolls of a biased die with where even numbers are three times as likely as odd numbers.

## Conditional statements
- if, else statement
- `ifelse` for vectorized operations

